---
title: "Social_Media_Raj_Motwani_project"
date: "2024-26-04"
output: html_document
---


```{r }
library(MASS)
library(ggplot2)
library(memisc)
library(ROCR)
library(dplyr)
library(klaR)
library(memisc)
library(ggplot2)
library(corrplot)
library(factoextra)
library(psych)
library(NbClust)
library(caTools)
library(magrittr)
library(pROC)
library(caret)


data <- read.csv("/Users/raj/downloads/Midterm_new.csv", row.names=1)
str(data)
```

DATA COLLECTION :
The dataset contains information on social media usage reported by 21 students, featuring 12 columns and 22 rows. Each row represents aggregated data for an individual student.


```{r }
str(data)
summary(data)
stars(data)
```

Data Dictionary :
Dataset ID: A unique identifier assigned to each student's data entry.
Instagram Usage: Duration of Instagram app usage.
LinkedIn Usage: Duration of LinkedIn app usage.
Snapchat Usage: Duration of Snapchat app usage.
Twitter Usage: Duration of Twitter app usage.
WhatsApp Usage: Duration of WhatsApp app usage.
YouTube Usage: Duration of YouTube app usage.
OTT Usage: Duration of Over-the-Top media services usage.
Reddit Usage: Duration of Reddit app usage.
Trouble Falling Asleep: Indicates whether the student reported difficulty falling asleep (0: No, 1: Yes).
Mood Productivity: Subjective measure of the student's mood and productivity level (0: Bad, 1: Good).
Tiredness upon Waking Up in the Morning: Indicates the level of tiredness reported by the student upon waking up in the morning (0: Low, 1: High).


```{r }
cor_max <- cor(data[,1:11])
cor_max

corrplot(cor(data), type = "upper", method = "color")

```

QUESTIONS and HYPOTHESIS Questions :
1. Can we determine if students experience tiredness upon waking up in the morning due to social media usage, using the provided information?
2. Can we predict whether students experience tiredness upon waking up in the morning due to social media usage, using the provided data?

Hypothesis :
We hypothesize that it is possible to utilize the duration of social media app usage by students to predict if they are affected by tiredness upon waking up in the morning.

The correlation matrix shows us a correlation between the columns in both cases. So, Principal Component Analysis (PCA) can help in reducing the number of columns used for analysis.


```{r }

# PCA

social_pca <- prcomp(data[,1:10],scale=TRUE) 
social_pca

summary(social_pca)

fviz_eig(social_pca, addlabels = TRUE)
```


The scree plot indicates that the total variance explained by the first 2 principal components is less than 70%. Therefore, using PCA to reduce the number of columns may not be suitable. Next, we'll explore Exploratory Factor Analysis (EFA) for this dataset.



```{r}

fviz_pca_var(social_pca,col.var = "cos2",
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"),
             repel = TRUE)
```

#### FACTOR ANALYSIS :

```{r}

fit.pc <- principal(data[,1:8], nfactors=4, rotate="varimax") 
fit.pc



fa.diagram(fit.pc) 

vss(data[,1:8])

```

RC1: Content Engagement: Measures user activity and interaction across content sharing platforms such as Twitter and Reddit.

RC2: Information Consumption: Reflects engagement with content consumption and learning activities, primarily through platforms like YouTube.

RC3: Communication and Entertainment: Represents usage patterns emphasizing communication and entertainment through platforms such as Whatsapp and OTT services.

RC4: Social Media Usage: Indicates the level of engagement with prominent social media platforms like Snapchat, LinkedIn, and Instagram.

CLUSTERING:

```{r}


efa_data <- as.data.frame(fit.pc$scores)
efa_data


matstd_sm <- scale(efa_data)



res.hc <- matstd_sm %>% scale() %>% dist(method = "euclidean") %>%
  hclust(method = "ward.D2")

res.nbclust <- matstd_sm %>% scale() %>% NbClust(distance = "euclidean", min.nc = 3, max.nc = 13, method = "complete", index ="all") 

km.res <- kmeans(matstd_sm, 5, nstart = 10)

fviz_cluster(km.res, data = matstd_sm,
             ellipse.type = "convex",
             palette = "jco",
             ggtheme = theme_minimal())

dist.data <- dist(matstd_sm, method="euclidean")



fviz_dend(res.hc, k = 5, # Cut in four groups
          cex = 0.5, # label size
          k_colors = c("#2E9FDF", "#00AFBB", "#E7B800", "#FC4E07"),
          color_labels_by_k = TRUE, # color labels by groups
          rect = TRUE # Add rectangle around groups
)


Clustered <- ifelse(km.res$cluster > 1.5, "NO Trouble waking up", "Trouble waking up")
Actual <- ifelse(data$Trouble_falling_asleep == 1, "Trouble waking up", "No trouble waking up ")
confusion_mat <- table(Clustered, Actual)
confusion_mat

accuracy <- sum(diag(confusion_mat)) / sum(confusion_mat)
precision <- confusion_mat[2, 2] / sum(confusion_mat[, 2])
recall <- confusion_mat[2, 2] / sum(confusion_mat[2, ])
cat("Accuracy:", round(accuracy, 3), "\n")


```


The confusion matrix indicates that the clustering predominantly categorizes users as not experiencing any difficulty falling asleep.The accuracy is 60%, which is not that good.


REGRESSION :

```{r}
wdbc <- read.csv("/Users/raj/downloads/Midterm_new.csv")
features <- c("Instagram", "LinkedIn", "Snapchat", "Twitter", "Whatsapp", "Youtube", "OTT", "Reddit", "Trouble_sleep", "Mood", "Tired_morning")
names(wdbc) <- c("ID", "Instagram", "LinkedIn", "Snapchat", "Twitter", "Whatsapp", "Youtube", "OTT", "Reddit", "Trouble_sleep", "Mood", "Tired_morning")
wdbc<- wdbc
wdbc.data <- as.matrix(wdbc[,c(2:9)])
row.names(wdbc.data) <- wdbc$ID

```


DATA SPLITTING :
```{r}
wdbc_raw <- cbind(wdbc.data, as.numeric(as.factor(wdbc$Trouble_sleep))-1)
colnames(wdbc_raw)[9] <- "Tired_waking_up_in_morning"
# Create train and test datasets
train_raw.df <- as.data.frame(wdbc_raw)
test_raw.df <- as.data.frame(wdbc_raw)
test_raw.df

```


LOGISTIC REGRESSION :


Logistic regression is a statistical approach used to classify data into two categories. It estimates the likelihood that an observation belongs to one of two classes. It establishes the connection between a binary outcome and one or multiple independent variables.

```{r}
data <- read.csv("/Users/raj/downloads/Midterm_new.csv")
logistic_data <- glm(Tired_waking_up_in_morning ~ ., data = data, family = 'binomial')
summary(logistic_data)

predicted.data <- data.frame(probability.of.hd=logistic_data$fitted.values,Tired_waking_up_in_morning=data$Tired_waking_up_in_morning)
predicted.data <- predicted.data[order(predicted.data$probability.of.hd, decreasing=FALSE),]
predicted.data$rank <- 1:nrow(predicted.data)

ggplot(data=predicted.data, aes(x=rank, y=probability.of.hd)) +
  geom_point(aes(color=Tired_waking_up_in_morning), alpha=1, shape=4, stroke=2) +
  xlab("Index") +
  ylab("Predicted probability of getting in sleeping")
probabilities_sm2 <- predict(logistic_data, new_data = test_raw.df, type = "response")
probabilities_sm2

predicted_sm2 <- ifelse(probabilities_sm2 > 0.5, "Yes", "No")
predicted_sm2
actual_sm <- ifelse(test_raw.df$Tired_waking_up_in_morning == 1, "Yes", "No")

confusion_sm2 <- table(predicted_sm2, actual_sm)
confusion_sm2
accuracy2 <- sum(diag(confusion_sm2)) / sum(confusion_sm2)
precision2 <- confusion_sm2[2, 2] / sum(confusion_sm2[, 2])
recall2 <- confusion_sm2[2, 2] / sum(confusion_sm2[2, ])
cat("Accuracy:", round(accuracy2, 3), "\n")
cat("Precision:", round(precision2, 3), "\n")
cat("Recall:", round(recall2, 3), "\n")
probabilities <- predict(logistic_data, new_data = test_raw.df, type = "response")

roc_sm <- roc(test_raw.df$Tired_waking_up_in_morning, probabilities_sm2)
auc_sm <- auc(roc_sm)

ggroc(roc_sm, color = "blue", legacy.axes = TRUE) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "False Positive Rate", y = "True Positive Rate",
       title = paste("ROC Curve (AUC = ", round(auc_sm, 2), ")")) +
  annotate("text", x = 0.5, y = 0.5, label = paste0("AUC = ", round(auc_sm, 2)))
```


CONCLUSION:
The logistic regression model exhibited remarkable accuracy and exceptional performance, as evidenced by its ROC curve. It notably excelled in accurately predicting outcomes, particularly in distinguishing between individuals experiencing sleep disturbances and those who are not, leveraging their social media usage patterns as predictive factors. This indicates the potential of utilizing the duration of time spent on various social media platforms to forecast whether a student is impacted by or encountering challenges in waking up early.